### 2013-10-14 Publications/2013-TMI/03-introduction.tex
Applying @zosso's changes to draft
Finished up to the numerical implementation section

\section{Introduction}
\label{sec:introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
%
\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
of \gls{mri} techniques \citep{sundgren_diffusion_2004} which 
accounts for a growing interest in its application to structural 
connectivity analyses of the brain at the macro-scale
\citep{craddock_imaging_2013}. This emerging field exploits
\gls{dwi} data to derive the local axonal structure at each imaging voxel 
\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
trajectories reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the 
\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
for the clinical application \citep{griffa_structural_2013} as well.

Early \gls{dwi} research mainly focused on the improvements of imaging 
methodologies by a better understanding of the diffusion effect, and improving
image reconstruction methodologies. Currently, the connectome extraction 
and analysis relies on a large chain of sophisticated computational techniques
including acquisition, reconstruction, modeling and model fitting, image segmentation
and registration, fibre tracking, connectivity mapping, visualization, statistical 
group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
has given rise to challenging issues towards reliable structural information about 
the neuronal tracts presented in, and also difficulties on the statistical analysis 
\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
regarding the impact of the different sources of inaccuracy and variability along
the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
cheng_characteristics_2012,jones_white_2013} are scarce.
Here, we shall address three tasks included within  the image processing stage in 
a unified approach: brain tissue segmentation in diffusion space 
(\autoref{sec:dwi_segmentation}), 
correction of geometrical distortions (\autoref{sec:distortion}), and structural 
image registration to diffusion coordinate space (\autoref{sec:registration}). 
These tasks are generally
solved independently, or combined in pairs. However, there exist fundamental 
coupling relationships that can be exploited to obtain a simultaneous solution to 
the three problems. This joint approach satisfactory impacts the downstream
outcomes of the whole pipeline with the increase of the internal consistency.

\subsection{\Gls{dwi} data overview}
\label{sec:dwi_overview}
\gls{dwi} data are usually acquired with \gls{epi} sequences 
as they allow for very fast acquisition of large sets of slices each
acquired after a single excitation \citep{mukherjee_diffusion_2008,
soares_hitchhikers_2013}. \Gls{epi} volumes are collected
in sequence and each one represents one probed diffusion direction.
This first acquisition is called \emph{raw} data or \emph{diffusion
signal}. One or more volumes without directional encoding (so-called \emph{baseline},
$b = 0$, \emph{low $b$}, or just \emph{B0} volumes) are interleaved as
reference, at the least one is acquired at the beginning of the
\gls{epi} dataset. These B0 are acquired without direction gradient,
showing a T2-like contrast.

@@

After correcting for artifacts, the difficulty of performing analyses directly
on the \emph{directionally dependent} raw data is not affordable. Thus, 
\gls{dwi} data is reconstructed using one of a numerous set of readily 
available algorithms. The outcome is a voxel-wise model of diffusion shape 
(in \citep{daducci_isbi2013_2013} \todo{FIX ref} there is available a comprehensive review of 
state-of-art methodologies). On top of the
chosen model, some scalar maps describing tissue features is generally
computed. For instance, the most popular \gls{dwi}-derived scalar map 
is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
which reflects the intensity of the diffusion effect. Many other
directionally invariant measurements have been defined. For a comprehensive study, 
we refer to \citep{ennis_orthogonal_2006}.

Alongside distortions and direction dependency, \gls{dwi} data is strongly 
affected by \glspl{pve} \citep{alexander_analysis_2001} that appears when
several different tissues, or signal 
emitters, are present in the same imaging unit, producing an averaged intensity. 
The effect is directly related to the low resolution achievable with \gls{dwi} 
(typically around $2.0\times2.0\times2.0mm^3$). An additional complication specific to
\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
respect \gls{csf}.

@@

A number of methods have been proposed using features not directly derived from \gls{dwi} data.
Segmentation obtained by co-registering structural T1-weighted images will be covered in 
\autoref{sec:registration}.

@@

Some other retrospective methodologies do not make use of the fieldmaps, 
as explicitly modeling the distortion \citep{andersson_modeling_2001},
registering with (anatomically correct) T2-weighted \gls{mri} 
\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
tao_variational_2009} (we shall cover these methods in 
\autoref{sec:registration}), or acquiring an 
extra B0 image with reversed phase encoding direction 
\citep{andersson_how_2003,holland_efficient_2010}.

@@

\subsection{Summary}
\label{sec:contributions}

@@

In this paper we propose a novel registration framework to simultaneously
solving the segmentation, distortion and cortical parcellation challenges, 
by exploiting as strong shape-prior the detailed morphology extracted 
from high-resolution and anatomically correct \gls{mri}. Indeed, hereafter 
we assume this segmentation problem in anatomical images is reliably and
accurately solved with readily available tools (e.g. \citep{fischl_freesurfer_2012}).
After global alignment with \gls{t1} using existing approaches, the remaining 
spatial mismatch between anatomical and diffusion space is due to susceptibility
distortions.
Finally, we need to establish precise spatial correspondence between the 
surfaces in both spaces, including the tangential direction for parcellation.
Therefore, we can reduce the problem to finding the differences of spatial 
distortion in between anatomical and \gls{dwi} space.
{\color{red} {We thus reformulate the segmentation problem as an inverse problem, where we 
seek for an underlying deformation field (the distortion) mapping 
from the structural space into the diffusion space, such that the structural 
contours segment optimally the \gls{dwi} data. In the process, the one-to-one 
correspondence between the contours in both spaces is guaranteed, and projection 
of parcellisation to \gls{dwi} space is implicit and consistent.

@@

The rest of this paper is organized as follows. First, in \autoref{sec:methods}
we introduce our proposed model for joint multivariate segmentation-registration.
Then we provide a more detailed description of the data and experimental setup in
\autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
in \autoref{sec:conclusion}.}}

>>>>

\section{Introduction}
\label{sec:introduction}

\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
of \gls{mri} techniques \citep{sundgren_diffusion_2004} which 
accounts for a growing interest in its application to structural 
connectivity analyses of the brain at the macro-scale
\citep{craddock_imaging_2013}. This emerging field exploits
\gls{dwi} data to derive the local axonal structure at each imaging voxel 
\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
trajectories reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the 
\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
for the clinical application \citep{griffa_structural_2013} as well.

% REMOVE paragraph (Dom, 1st comment)
%Early \gls{dwi} research mainly focused on the improvements of imaging 
%methodologies by a better understanding of the diffusion effect, and improving
%image reconstruction methodologies. Currently, the connectome extraction 
%and analysis relies on a large chain of sophisticated computational techniques
%including acquisition, reconstruction, modeling and model fitting, image segmentation
%and registration, fibre tracking, connectivity mapping, visualization, statistical 
%group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
%has given rise to challenging issues towards reliable structural information about 
%the neuronal tracts presented in, and also difficulties on the statistical analysis 
%\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
%regarding the impact of the different sources of inaccuracy and variability along
%the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
%cheng_characteristics_2012,jones_white_2013} are scarce.

\subsection{\Gls{dwi} data overview}
\label{sec:dwi_overview}
\gls{dwi} data are usually acquired with \gls{epi} schemes
as they allow for very fast acquisition of large sets of slices each
acquired after a single excitation \citep{mukherjee_diffusion_2008,
soares_hitchhikers_2013}. \Gls{epi} volumes are collected
in sequence and each one represents one probed diffusion direction.
%This first acquisition is called \emph{raw} data or \emph{diffusion
%signal}. One or more volumes without directional encoding (so-called \emph{baseline},
%$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
%reference, at the least one is acquired at the beginning of the
%\gls{epi} dataset. These B0 are acquired without direction gradient,
%showing a T2-like contrast.

@@

After correcting for artifacts, the difficulty of performing analyses directly
on the \emph{directionally dependent} raw data is not affordable. Thus, 
\gls{dwi} data is reconstructed using one of a numerous set of readily 
available algorithms. The outcome is a voxel-wise model of diffusion shape 
(in \citep{daducci_isbi2013_2013} \todo{FIX ref} there is available a comprehensive review of 
state-of-art methodologies). On top of the
chosen model, some scalar maps describing tissue features are generally
computed (e.g. \gls*{fa}, and \gls*{adc}).
%For instance, the most popular \gls{dwi}-derived scalar map 
%is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
%Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
%which reflects the intensity of the diffusion effect. Many other
%directionally invariant measurements have been defined. For a comprehensive study, 
%we refer to \citep{ennis_orthogonal_2006}.

Alongside distortions and direction dependency, \gls{dwi} data is strongly 
affected by \glspl{pve} \citep{alexander_analysis_2001} that appears when
several different tissues are present in the same imaging unit, producing
an averaged intensity. The effect is directly related to the resolution,
typically around $2.0\times2.0\times2.0mm^3$ for \gls{dwi}.

% the low resolution achievable with \gls{dwi} 
%(typically around ). An additional complication specific to
%\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
%that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
%linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
%respect \gls{csf}.

Some other retrospective methodologies do not make use of the fieldmaps, 
as explicitly modeling the distortion \citep{andersson_modeling_2001},
registering with (anatomically correct) T2-weighted \gls{mri} 
\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
tao_variational_2009}, or acquiring an 
extra B0 image with reversed phase encoding direction 
\citep{andersson_how_2003,holland_efficient_2010}.

@@

%A number of methods have been proposed using features not directly derived from \gls{dwi} data.
%Segmentation obtained by co-registering structural T1-weighted images will be covered in 
%\autoref{sec:registration}.

@@

@@

\subsection{Proposed model}
\label{sec:contributions}

@@

% Paragraph taken from first subsection to here
Here, we shall address three tasks included within  the image processing stage in 
a unified approach: brain tissue segmentation in diffusion space 
(\autoref{sec:dwi_segmentation}), 
correction of geometrical distortions (\autoref{sec:distortion}), and structural 
image registration to diffusion coordinate space (\autoref{sec:registration}). 
These tasks are generally
solved independently, or combined in pairs. However, there exist fundamental 
coupling relationships that can be exploited to obtain a simultaneous solution to 
the three problems. This joint approach satisfactory impacts the downstream
outcomes of the whole pipeline with the increase of the internal consistency.

In this paper we propose a novel registration framework to simultaneously
solving the segmentation, distortion and cortical parcellation challenges, 
by exploiting as strong shape-prior the detailed morphology extracted 
from high-resolution and anatomically correct \gls{mri}. Indeed, hereafter 
we assume this segmentation problem in anatomical images is reliably and
accurately solved with readily available tools (e.g. \citep{fischl_freesurfer_2012}).
After global alignment with \gls{t1} using existing approaches, the remaining 
spatial mismatch between anatomical and diffusion space is due to susceptibility
distortions.
Finally, we need to establish precise spatial correspondence between the 
surfaces in both spaces, including the tangential direction for parcellation.
Therefore, we can reduce the problem to finding the differences of spatial 
distortion in between anatomical and \gls{dwi} space.
We thus reformulate the segmentation problem as an inverse problem, where we 
seek for an underlying deformation field (the distortion) mapping 
from the structural space into the diffusion space, such that the structural 
contours segment optimally the \gls{dwi} data. In the process, the one-to-one 
correspondence between the contours in both spaces is guaranteed, and projection 
of parcellisation to \gls{dwi} space is implicit and consistent.

@@

The rest of this paper is organized as follows. First, in \autoref{sec:methods}
we introduce our proposed model for joint multivariate segmentation-registration.
Then we provide a more detailed description of the data and experimental setup in
\autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
in \autoref{sec:conclusion}.
